The impact of gas bulk rotation on the Lyman $\alpha$ line

### 2014-06-04 paper/referee-reply.tex

We do not find any significant dependence of the escape fraction with
the viewing angle. This is shown in Figure \ref{escape_fraction} of this
document. We do not show this plot in the paper and only mention the
result at the beginning of Section 3.5.

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{escapefractionvstheta.png}
\end{center}
\caption{Escape Fraction as a function of $\mu$
\label{escape_fraction}}
\end{figure}

\section*{Outflow + Rotation}

{\bf So far, the model only includes rotation. However, most observed
  Lya spectra from LAEs, or even local galaxies, seem to be asymetric
  towards the red wavelengths, interpreted as a sign for outflows in
  these systems, often corroborated by others observables. Did you
  investigate how the rotation would modify the spectra emerging from
  expanding clouds?}

We agree with the referee that this is an important point to study. 
We have not fully investigated it yet.  This is actually research
under development.

@@

Done.

\subsection*{Fig.1}

{\bf I guess that the spectra presented in Fig1 are integrated over
  all directions, right ? You should describe explicitly how you build
  them. You could skip the x notation in absciss, as it is not used in
  the discussion, whereas the velocity is used to compare to FWHM, on
  Fig2.}

Now this figure is the Fig.4 in the paper. Given the depence on the
viewing angle we decided to construct the plots for a line-of-sight
perpendicular to the rotation axis ($\theta=\pi/2$). This is
explicitly describe in the figure and the caption. We have also
changed the $x$ notation to velocities.

@@

\subsection*{Fig.2}

{\bf Fig.2 Can you explain how you measure the FWHM of a double-peaked
  profile ? Do you fit it with a gaussian ?}

We do not make a gaussian fit. We build first a histogram for the
outgoing frequencies. Then we interpolate between the points of the
histogram to find the half maxima intensity values and their
corresponding frequencies.

@@

\subsection*{Fig.3}

{\bf Do you have an idea why the (central source, intermediate optical
  depth) case with Vmax=300 has a single plateau instead of 2 peaks ?
  Do you find this with the two codes ?}

After close inspection if this figure the line was composed by two 
peaks that looked like a plateau with the resolution used to
construct the figure.

However, this plateau is not present in Fig. 4 anymore because we do
not build the line from all the outgoing photons, but instead take
into account the line of sight for the observation.

@@

(MARK)

\subsection*{Fig5}

{\bf This is a very nice figure ! Looking at the top right panel, with
  its “photosphere”, I’m surprised that the distribution of Nscatt is
  bipolar, I would have expected a smooth transition between the 2
  regimes. Do you have a physical explanation why photons escape after
  either (less than 10) or (more than 1000) scatterings, and none
  escape with 100 scatt ? To test the ’photosphere’ assumption, you
  could also do the radiation transfer in a cloud with the
  distribution of sources only in an inner sphere, and another
  experienment with sources only in an outer shell. If I understood
  corectly, you would expect thant the bottom spot on the top right
  panel is made of photons from the outer shell, and the uppper spot
  from photons emitted inside the cloud.}

We followed the suggestion by the referee to construct a new
plot. For the model with homogeneously distributed sources we now the
initial and final states for every photon. With this information we
prepare the plot in Figure \ref{nscatt} of this document. It is a 2D
histogram of the number of scatterings and the initial radius showing
the starting radius. Indeed, most of the photons that escaped after
less than 10 scatterings were emitted in the external part of the
sphere $R>0.8R_{max}$. However, there is a minority of photons that
have a similar emission radius but only managed to escape after $10^4$
scatterings.

\begin{figure}
\begin{center}
  \includegraphics[scale=0.6]{Histogram2dNscattVSRadius.png}
\end{center}
\caption{\label{nscatt} 2D histogram of the number of scatterings and
  the number. The color scale is linear in the number of photons in
  each bin.}
\end{figure}

@@

(MARK)

\subsection*{Fig8}

{\bf Referred to as Figure 7 in the text (paragraph 3.4). To my mind,
  this figure illustrates the main result of your study, it has to be
  more demonstrative. On this figure, the numerical noise seems very
  big compare to the small number of bins, bigger than on the spectra
  from Fig1, how is it possible ? Could you re-do this figure with
  more photons, maybe with more bins ? Could you plot an histogram
  instead of broken lines, to ease the reading ? In the text, you
  write anisotropy induced by rotation is at the 3$\%$ level, how do
  to estimate this number? I think that the noise is at $3\%$ level,
  and the distribution is compatible with ’flat’. In the corresponding
  text, you first say that $F(\mu)$ does not depend on Vmax. Again, if
  real, this is an important result, that you must advertise
  more. Please show it on this Fig, by comparing F($\mu$) without rotation
  (in red, for example) to the other distributions that you get with
  different values of Vmax. Then, you mention that for high optical
  depth values, $F(\mu)$ (not the variations of $\mu$) can vary up to
  $15\%$. First, what do you call high optical depth ? your highest
  setup is $\tau$= $10^7$, not extremely high for lyman alpha ”standards’,
  corresponding to NHI $\sim$10$^{20}$ cm$^{-2}$. So, if you see a trend with
  optical depth, I encourage you to check higher optical depth
  regimes, corresponding better to galaxy scales. If you can probe
  that anisotropy induced by rotation is rising with optical depth,
  this would be an interesting result. If higher optical depth regimes
  lead to anisotropic escape, could you check the effet of anisotropy
  on the lya spectral shape?}

We have prepared Figure 5 that shows the $F(\mu)$ distribution for
different velocities and optical depths. Indeed, the distributions are
consistent with flat (within the noise levels) for all the models
studied in our paper.

We did not try many more models with higher optical depths. We expect
to do that in the near future to complete a study that could be
immediatly comparable against observations.

### 2014-06-04 paper/ms.tex
single flight clarification

In the right panel of Figure \ref{fig:escape_fraction} we put these
results in the context of the analytic solution for the infinite slab
\citep{Neufeld90}. In Neufeld's set-up the analytic solution depends
uniquely on the product $(a\tau_{\rm   H})^{1/3}\tau_{A}$ where
$\tau_{A} = (1 - A)\tau_{a}$, valid only in the limit $a\tau_{\rm
  H}\gg 1$. The dashed lines in  the right panel of Figure
\ref{fig:escape_fraction} show the results for the different
rotational velocities for the homogeneous models. First of all we note
that the escape fraction does not increase significantly from
$\tau_{\rm H}=10^{5}$ to $\tau_{\rm H}=10^{6}$. This
counter-intuitive result is a consequence of the transition into the
'extremely' opaque regime,  which occurs roughly when $a\tau_{\rm{H}}
> 10^{3}$  e.g. \citep{Neufeld90}.

>>>>

In the right panel of Figure \ref{fig:escape_fraction} we put these
results in the context of the analytic solution for the infinite slab
\citep{Neufeld90}. In Neufeld's set-up the analytic solution depends
uniquely on the product $(a\tau_{\rm   H})^{1/3}\tau_{A}$ where
$\tau_{A} = (1 - A)\tau_{a}$, valid only in the limit $a\tau_{\rm
  H}\gg 1$. The dashed lines in  the right panel of Figure
\ref{fig:escape_fraction} show the results for the different
rotational velocities for the homogeneous models. First of all we note
that the escape fraction increases slightly from $\tau_{\rm H}=10^5$ to
$\tau_{\rm H}=10^{6}$, which is counterintuitive as we expected that increasing
$\tau_{\rm H}$ requires more scattering before the photon escapes, which would
increase the probability that photons are destroyed by dust.

However, this naive interpretation is incorrect: increasing $\tau_{\rm
  H}$ from $10^{5}$ to $10^{6}$ causes us to transition from the
'optically thick' to  the 'extremely optically thick'
  regime. In the optically thick regime, Lya  photons can escape in
  'single flight' which corresponds to a scenario in which the
  photon resonantly scatters $10^4-10^5$ times until it is scattered
  into the wing of the line ($x\sim 3-4$). At these frequencies the
  medium is optically thin, and the photons can escape efficiently in
  a single flight. In contrast, in an extremely optically thick medium \ly
  photons escape in a `single excursion' \citep{Adams72}. Here, photons
that are scattered into the wing of the line escape from the medium in
a sequence of wing scattering events.In both cases, \ly photons
resonantly scatter $10^4$-$10^5$ times. Because we keep our clouds the same
size, the mean free path of Lya photons that scatter resonantly is 10
times larger for the case $\tau_{\rm H}=10^5$ than for $\tau_{\rm
  H}=10^6$. If we compute the average distance $D$ travelled by \ly
photons through a medium of size $R$ as a function of line center
optical depth $\tau_{\rm H}$, then we find that during the transition
from optically thick to extremely optically thick the mean traversed
distance $D$ actually decreases slightly.  This decrease is unique to
this transition region, and $D$ generally increases with $\tau_{\rm H}$ at other
values of $\tau_{\rm H}$.
